---
title: "Alfalfa Analysis"
author: "Nick Callen, Dela Hatfield, Cole Ternes, Logan Welsh"
date: "3/28/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import_data, echo=FALSE}
#Imports data
setwd("C:/Users/Cole Ternes/Desktop/CPSC_courses/personalprojects/Alfalfa")
data.raw <- read.csv("raw_data.csv", stringsAsFactors = F)
#Call libraries
library(ggplot2)
```

```{r prep_data, echo=FALSE}
#Vectors
n_sa_rep <- c()
n_sa_nodes <- c()
n_sa_mass <- c()

n_te_rep <- c()
n_te_nodes <- c()
n_te_mass <- c()

n_ma_rep <- c()
n_ma_nodes <- c()
n_ma_mass <- c()

y_sa_rep <- c()
y_sa_nodes <- c()
y_sa_mass <- c()

y_te_rep <- c()
y_te_nodes <- c()
y_te_mass <- c()

y_ma_rep <- c()
y_ma_nodes <- c()
y_ma_mass <- c()

#For-Loop to input data into vectors
for (i in 1:nrow(data.raw)) {
  if (data.raw[i,1] == "N_Sa") {
    n_sa_rep <- c(n_sa_rep, data.raw[i,4])
    n_sa_nodes <- c(n_sa_nodes, data.raw[i,5])
    n_sa_mass <- c(n_sa_mass, data.raw[i,6])
  } else if (data.raw[i,1] == "N_Te") {
    n_te_rep <- c(n_te_rep, data.raw[i,4])
    n_te_nodes <- c(n_te_nodes, data.raw[i,5])
    n_te_mass <- c(n_te_mass, data.raw[i,6])
  } else if (data.raw[i,1] == "N_Ma") {
    n_ma_rep <- c(n_ma_rep, data.raw[i,4])
    n_ma_nodes <- c(n_ma_nodes, data.raw[i,5])
    n_ma_mass <- c(n_ma_mass, data.raw[i,6])
  } else if (data.raw[i,1] == "Y_Sa") {
    y_sa_rep <- c(y_sa_rep, data.raw[i,4])
    y_sa_nodes <- c(y_sa_nodes, data.raw[i,5])
    y_sa_mass <- c(y_sa_mass, data.raw[i,6])
  } else if (data.raw[i,1] == "Y_Te") {
    y_te_rep <- c(y_te_rep, data.raw[i,4])
    y_te_nodes <- c(y_te_nodes, data.raw[i,5])
    y_te_mass <- c(y_te_mass, data.raw[i,6])
  } else if (data.raw[i,1] == "Y_Ma") {
    y_ma_rep <- c(y_ma_rep, data.raw[i,4])
    y_ma_nodes <- c(y_ma_nodes, data.raw[i,5])
    y_ma_mass <- c(y_ma_mass, data.raw[i,6])
  } 
}
```


```{r plot_n_nodes_df, echo=FALSE}
#No Bacteria Node_Plot Variables
n_sa_nodes_df <- data.frame(x=n_sa_rep, y=n_sa_nodes)
n_te_nodes_df <- data.frame(x=n_te_rep, y=n_te_nodes)
n_ma_nodes_df <- data.frame(x=n_ma_rep, y=n_ma_nodes)

n_sa_nodes_df$Treatment="Sand"
n_te_nodes_df$Treatment="Terrain"
n_ma_nodes_df$Treatment="Martian"

n_nodes_df=rbind(n_sa_nodes_df, n_te_nodes_df, n_ma_nodes_df)

n_sa_nodes_mean <- mean(n_sa_nodes)
n_te_nodes_mean <- mean(n_te_nodes)
n_ma_nodes_mean <- mean(n_ma_nodes)

#Plot
ggplot(n_nodes_df) + geom_line(aes(x, y, colour=Treatment)) + xlab("Replicates") + ylab("Number of Nodes")  + theme_minimal() + ggtitle("Number of Nodes per Replicate without Bacteria")
```

Mean Number of Nodes for Sand Treatment: `r n_sa_nodes_mean`

Mean Number of Nodes for Terrain Treatment: `r n_te_nodes_mean`

Mean Number of Nodes for Martian Treatment: `r n_ma_nodes_mean`

```{r plot_y_nodes_df, echo=FALSE}
#Yes Bacteria Node_Plot Variables
y_sa_nodes_df <- data.frame(x=y_sa_rep, y=y_sa_nodes)
y_te_nodes_df <- data.frame(x=y_te_rep, y=y_te_nodes)
y_ma_nodes_df <- data.frame(x=y_ma_rep, y=y_ma_nodes)

y_sa_nodes_df$Treatment="Sand"
y_te_nodes_df$Treatment="Terrain"
y_ma_nodes_df$Treatment="Martian"

y_nodes_df=rbind(y_sa_nodes_df, y_te_nodes_df, y_ma_nodes_df)

y_sa_nodes_mean <- mean(y_sa_nodes)
y_te_nodes_mean <- mean(y_te_nodes)
y_ma_nodes_mean <- mean(y_ma_nodes)

#Plot
ggplot(y_nodes_df) + geom_line(aes(x, y, colour=Treatment)) + xlab("Replicates") + ylab("Number of Nodes")  + theme_minimal() + ggtitle("Number of Nodes per Replicate with Bacteria")
```

Mean Number of Nodes for Sand Treatment: `r y_sa_nodes_mean`

Mean Number of Nodes for Terrain Treatment: `r y_te_nodes_mean`

Mean Number of Nodes for Martian Treatment: `r y_ma_nodes_mean`

``` {r plot_total_nodes, echo=FALSE}
#Total_Node_Plot Variables
n_sa_total_nodes <- sum(n_sa_nodes)
n_te_total_nodes <- sum(n_te_nodes)
n_ma_total_nodes <- sum(n_ma_nodes)
y_sa_total_nodes <- sum(y_sa_nodes)
y_te_total_nodes <- sum(y_te_nodes)
y_ma_total_nodes <- sum(y_ma_nodes)

total_nodes_df <- data.frame(x=c("No Bac/Sand", "No Bac/Terrain", "No Bac/Martian", "Yes Bac/Sand", "Yes Bac/Terrain", "Yes Bac/Martian"), y=c(n_sa_total_nodes, n_te_total_nodes, n_ma_total_nodes, y_sa_total_nodes, y_te_total_nodes, y_ma_total_nodes))

#Plot
ggplot(data=total_nodes_df, aes(x=x, y=y)) + geom_bar(stat="identity", fill="steelblue") + xlab("") + ylab("Total Number of Nodes") + theme_minimal() + ggtitle("Total Number of Nodes per Treatment") + geom_text(aes(label=y), vjust=-0.3, size=3.5)
```


``` {r plot_n_mass_df, echo=FALSE}
#No Bacteria Mass_Plot Variables
n_sa_mass_df <- data.frame(x=n_sa_rep, y=n_sa_mass)
n_te_mass_df <- data.frame(x=n_te_rep, y=n_te_mass)
n_ma_mass_df <- data.frame(x=n_ma_rep, y=n_ma_mass)

n_sa_mass_df$Treatment="Sand"
n_te_mass_df$Treatment="Terrain"
n_ma_mass_df$Treatment="Martian"

n_mass_df=rbind(n_sa_mass_df, n_te_mass_df, n_ma_mass_df)

n_sa_mass_mean <- mean(n_sa_mass)
n_te_mass_mean <- mean(n_te_mass)
n_ma_mass_mean <- mean(n_ma_mass)

#Plot
ggplot(n_mass_df) + geom_line(aes(x, y, colour=Treatment)) + xlab("Replicates") + ylab("Shoot Mass (mg)")  + theme_minimal() + ggtitle("Mass per Replicate without Bacteria")
```

Mean Shoot Mass for Sand Treatment: `r n_sa_mass_mean` mg

Mean Shoot Mass for Terrain Treatment: `r n_te_mass_mean` mg

Mean Shoot Mass for Martian Treatment: `r n_ma_mass_mean` mg

``` {r plot_y_mass_df, echo=FALSE}
#Yes Bacteria Mass_Plot Variables
y_sa_mass_df <- data.frame(x=y_sa_rep, y=y_sa_mass)
y_te_mass_df <- data.frame(x=y_te_rep, y=y_te_mass)
y_ma_mass_df <- data.frame(x=y_ma_rep, y=y_ma_mass)

y_sa_mass_df$Treatment="Sand"
y_te_mass_df$Treatment="Terrain"
y_ma_mass_df$Treatment="Martian"

y_mass_df=rbind(y_sa_mass_df, y_te_mass_df, y_ma_mass_df)

y_sa_mass_mean <- mean(y_sa_mass)
y_te_mass_mean <- mean(y_te_mass)
y_ma_mass_mean <- mean(y_ma_mass)

#Plot
ggplot(y_mass_df) + geom_line(aes(x, y, colour=Treatment)) + xlab("Replicates") + ylab("Shoot Mass (mg)")  + theme_minimal() + ggtitle("Mass per Replicate with Bacteria")
```

Mean Shoot Mass for Sand Treatment: `r y_sa_mass_mean` mg

Mean Shoot Mass for Terrain Treatment: `r y_te_mass_mean` mg

Mean Shoot Mass for Martian Treatment: `r y_ma_mass_mean` mg

``` {r plot_total_mass, echo=FALSE}
#Total_Mass_Plot Variables
n_sa_total_mass <- sum(n_sa_mass)
n_te_total_mass <- sum(n_te_mass)
n_ma_total_mass <- sum(n_ma_mass)
y_sa_total_mass <- sum(y_sa_mass)
y_te_total_mass <- sum(y_te_mass)
y_ma_total_mass <- sum(y_ma_mass)

total_mass_df <- data.frame(x=c("No Bac/Sand", "No Bac/Terrain", "No Bac/Martian", "Yes Bac/Sand", "Yes Bac/Terrain", "Yes Bac/Martian"), y=c(n_sa_total_mass, n_te_total_mass, n_ma_total_mass, y_sa_total_mass, y_te_total_mass, y_ma_total_mass))

#Plot
ggplot(data=total_mass_df, aes(x=x, y=y)) + geom_bar(stat="identity", fill="steelblue") + xlab("") + ylab("Total Shoot Mass (mg)") + theme_minimal() + ggtitle("Total Shoot Mass (mg) per Treatment") + geom_text(aes(label=y), vjust=-0.3, size=3.5)
```